---
title: "spread vs cluster"
author: "DC"
date: "6/22/2020"
output: html_document
---

#subset the hahn set into 5,5,<1 
#using regexes to find cluster + spread

```{r}
library(tidyverse)
library(stringr)
```
```{r}
#this is the data for the graph 
hahn_tads$Basicstotal<- str_count(hahn_tads$sequence, "[HRK]")
hahn_tads$Acidstotal<- str_count(hahn_tads$sequence, "[ED]")
hahn_tads$Aromaticstotal<- str_count(hahn_tads$sequence, "[WYF]")
clustertest_2<-hahn_tads %>% filter(Basicstotal<=1& Acidstotal==5 & Aromaticstotal==5)
```


```{r}
aro_1<-str_count(clustertest_2$sequence,"[WYF][WYF][WYF][WYF]")
aro_2<-str_count(clustertest_2$sequence,"[WYF][WYF][WYF][^WYF][WYF]")
aro_3<-str_count(clustertest_2$sequence,"[WYF][WYF][^WYF][WYF][WYF]")
aro_4<-str_count(clustertest_2$sequence,"[WYF][^WYF][WYF][WYF][WYF]")


arocount1<-which(aro_1>0)
hold1<-clustertest_2[arocount1,]

arocount2<-which(aro_2>0)
hold2<-clustertest_2[arocount2,]

arocount3<-which(aro_3>0)
hold3<-clustertest_2[arocount3,]

arocount4<-which(aro_4>0)
hold4<-clustertest_2[arocount4,]


arototal<-c(arocount1,arocount2,arocount3,arocount4)
#9/40 22.5%


table(clustertest_2[arototal,]$AD_set)

Kadabra<-unique(arototal)
table(clustertest_2[Kadabra,]$AD_set)
Starmie<-clustertest_2[Kadabra,] %>% filter(AD_set=="AD_positive")
sum(Starmie$enrichmen_score)/length(Starmie$enrichmen_score)

#new aro cluster 5.874444
```




```{r}
acid_1<-str_count(clustertest_2$sequence, "[DE][DE][DE][DE]")
acid_2<-str_count(clustertest_2$sequence,"[DE][DE][DE][^DE][DE]")
acid_3<-str_count(clustertest_2$sequence,"[DE][DE][^DE][DE][DE]")
acid_4<-str_count(clustertest_2$sequence,"[DE][^DE][DE][DE][DE]")


acidcount1<-which(acid_1>0)
hold1<-clustertest_2[acidcount1,]


acidcount2<-which(acid_2>0)
hold2<-clustertest_2[acidcount2,]

acidcount3<-which(acid_3>0)
hold3<-clustertest_2[acidcount3,]


acidcount4<-which(acid_4>0)
hold4<-clustertest_2[acidcount4,]

acidtotal<-c(acidcount1,acidcount2,acidcount3,acidcount4)

Abra<-unique(acidtotal)
table(clustertest_2[Abra,]$AD_set)

#enrichmen_score
Staryu<-clustertest_2[Abra,] %>% filter(AD_set=="AD_positive")
sum(Staryu$enrichmen_score)/length(Staryu$enrichmen_score)


#old score sum(clustertest_2[Abra,]$enrichmen_score)/length(clustertest_2[Abra,]$enrichmen_score)
 
#3/38 07%
#new score 3.85667

```

```{r}
table(clustertest_2$AD_set)
Magikarp<-clustertest_2 %>% filter(AD_set=="AD_positive")
sum(Magikarp$enrichmen_score)/length(Magikarp$enrichmen_score)


sum(clustertest_2$enrichmen_score)/length(clustertest_2$enrichmen_score)


#base case 737/2122 34.7%
#new score 4.975
```





```{r}
arospread_1<-str_count(clustertest_2$sequence,"[WYF][^WYF][^WYF][^WYF]")                      
aro_spreadcount1<-which(arospread_1 ==5)
hold1<-clustertest_2[aro_spreadcount1,]

arospread_2<-str_count(clustertest_2$sequence,"[^WYF][^WYF][^WYF][WYF]")                    
aro_spreadcount2<-which(arospread_2 ==5)
hold2<-clustertest_2[aro_spreadcount2,]


Hocus<-unique(c(aro_spreadcount1,aro_spreadcount2))
table(clustertest_2[Hocus,]$AD_set)

#total 35/84 41.6

#enrichmen_score
Horsea<-clustertest_2[Hocus,] %>% filter(AD_set=="AD_positive")
sum(Horsea$enrichmen_score)/length(Horsea$enrichmen_score)
#old sum(clustertest_2[Hocus,]$enrichmen_score)/length(clustertest_2[Hocus,]$enrichmen_score)
#new score 4.983
```



```{r}
#use this one 
acidspread_1<-str_count(clustertest_2$sequence,"[DE][^DE][^DE][^DE]")                      
acid_spreadcount1<-which(acidspread_1 ==5)
hold1<-clustertest_2[acid_spreadcount1,]

acidspread_2<-str_count(clustertest_2$sequence,"[^DE][^DE][^DE][DE]")                      
acid_spreadcount2<-which(acidspread_2 ==5)
hold2<-clustertest_2[acid_spreadcount2,]

Pocus<-unique(c(acid_spreadcount1,acid_spreadcount2))
table(clustertest_2[Pocus,]$AD_set)
#40/95 43.4%
#enrichmen_score
Seadra<-clustertest_2[Pocus,] %>% filter(AD_set=="AD_positive")
sum(Seadra$enrichmen_score)/length(Seadra$enrichmen_score)
#old sum(clustertest_2[Pocus,]$enrichmen_score)/length(clustertest_2[Pocus,]$enrichmen_score)
# new score 4.983


```
```{r}

```

```{r}
#Numbers at the top of the bars represent total number of sequences in that given pool & numbers in the middle of the bars represent the average enrichmen_score scores for that pool
```
                                      

```{r}
#only positive enrichmen_score 
average_enirchment_scores<-c(4.975,4.83,3.85667,4.983,5.810769)
data<-data.frame(case=c("Base Case", "Acid Spread", "Acid Clustered", "Aro Spread", "Aro Clustered"), live_percentage=c(34.7, 43.4,7,41.6, 22.5))
sequence_count<- c(2122,92,38,84,40)
data$enrichment_scores=average_enirchment_scores
as.factor(data$enrichment_scores)
ggplot(data, aes(x=case , y=live_percentage,fill=enrichment_scores))+
geom_bar(stat="identity")+scale_fill_gradient2(low="blue",high="red",mid = "grey",midpoint = 4.975)+labs(x=" ", y="live percentage",title = "Analysis of clustering vs spread of acidic and aromatic amino acids")+ geom_text(aes(label=sequence_count), vjust=-0.3, size=5)+  scale_y_continuous(breaks = scales::pretty_breaks(n = 10)) +theme_classic() 




```

