---
title: "Machine Learning Comparison"
output: html_notebook
---


```{r}

dat <- read.csv("../data/xiao_analysis/Cmp_Attention_Ridge.csv",
                stringsAsFactors = FALSE)

pn.stats <- table(dat$AD_label)

```

This analysis includes `r nrow(dat)` sequences from the test set (consistent for
both ML methods), of which `r sum(dat$AD_label)` were AD_positive.

# Within-method analysis

```{r, fig.height = 5, fig.width = 5}

library(ROCR)

pred.ridge <- prediction(predictions = dat$Ridge_positive_prob, 
                          labels = dat$AD_label)
perf.ridge <- performance(pred.ridge, measure = "tpr", x.measure = "fpr")
plot(perf.ridge, col=rainbow(10))

```
We'll select the cutoff for this analysis based on the cutoff that maximizes
(TPR + TNR).

```{r}
pred.ridge.table <- data.frame(cutoff = pred.ridge@cutoffs[[1]],
                               tpr = pred.ridge@tp[[1]] / pn.stats[2],
                               tnr = pred.ridge@tn[[1]] / pn.stats[1])
pred.ridge.table$ind <- pred.ridge.table$tpr + pred.ridge.table$tnr - 1

best.cutoff.ind <- which.max(pred.ridge.table$ind)
best.cutoff.ridge <- round(pred.ridge.table$cutoff[best.cutoff.ind], 3)
pred.ridge.table[(best.cutoff.ind-10):(best.cutoff.ind+10),]

```

The best cutoff is `r best.cutoff.ridge`.

```{r, fig.height = 5, fig.width = 5, eval = FALSE}

perf.ridge <- performance(pred.ridge, "cal")
plot(perf.ridge, col=rainbow(10))

```


```{r, fig.height = 5, fig.width = 5}

pred.nn <- prediction(predictions = dat$Attention_positive_prob, 
                          labels = dat$AD_label)
perf.nn <- performance(pred.nn, measure = "tpr", x.measure = "fpr")
plot(perf.nn, col=rainbow(10))

```

```{r}
pred.nn.table <- data.frame(cutoff = pred.nn@cutoffs[[1]],
                               tpr = pred.nn@tp[[1]] / pn.stats[2],
                               tnr = pred.nn@tn[[1]] / pn.stats[1])
pred.nn.table$ind <- pred.nn.table$tpr + pred.nn.table$tnr - 1

best.cutoff.ind <- which.max(pred.nn.table$ind)
best.cutoff.nn <- round(pred.nn.table$cutoff[best.cutoff.ind], 2)
pred.nn.table[(best.cutoff.ind-10):(best.cutoff.ind+10),]

```

The best cutoff is `r best.cutoff.nn`.


# Method comparison

Overall, the methods are relatively concordant, with a Spearman correlation of
`r cor(dat$Ridge_positive_prob, dat$Attention_positive_prob, method = "spearman")`

For the positive sequences, we see the following prediction accuracy from the 
two methods, by using the optimal cutoffs defined above.

```{r}

dat.pos <- dat[dat$AD_label == 1,]
dat.pos$combined_prob <- dat.pos$Ridge_positive_prob + dat.pos$Attention_positive_prob
dat.pos$diff_prob <- dat.pos$Ridge_positive_prob - dat.pos$Attention_positive_prob

table(dat.pos$Ridge_positive_prob > best.cutoff.ridge,
      dat.pos$Attention_positive_prob > best.cutoff.nn)

```
Some False-Negatives identified by Ridge regression were borderline cases,
while others had very low prediction values.

```{r}
hist(dat.pos$Ridge_positive_prob[dat.pos$Ridge_positive_prob < best.cutoff.ridge])
```
This is similar for the neural network approach, although many of them had very
low probabilities

```{r}
hist(dat.pos$Attention_positive_prob[dat.pos$Attention_positive_prob < best.cutoff.nn])
```


For the negative sequences, we see the following prediction accuracy from the 
two methods.

```{r}

dat.neg <- dat[dat$AD_label == 0,]

dat.neg$combined_prob <- dat.neg$Ridge_positive_prob + dat.neg$Attention_positive_prob
dat.neg$diff_prob <- dat.neg$Ridge_positive_prob - dat.neg$Attention_positive_prob


table(dat.neg$Ridge_positive_prob < best.cutoff.ridge,
      dat.neg$Attention_positive_prob < best.cutoff.nn)

```

Using the optimal cutoffs, we clearly see better performance by the attention
model, and it's relatively better in correctly identifying negative cases.

Many False-Positives identified by Ridge regression were borderline cases.

```{r}
hist(dat.neg$Ridge_positive_prob[dat.neg$Ridge_positive_prob > best.cutoff.ridge],
     breaks = seq(0.04, 1, by=0.05))
```

Many False-Positives identified by the attention model had high probabilities.

```{r}
hist(dat.neg$Attention_positive_prob[dat.neg$Attention_positive_prob > best.cutoff.nn])
```

## Scatterplot

Predictions for positive sequences

```{r, fig.height = 5, fig.width = 7}

library(ggplot2)

ggplot(dat.pos, aes(x = Ridge_positive_prob, y = Attention_positive_prob)) +
  geom_point(alpha = 0.05) + geom_hline(yintercept = best.cutoff.nn) +
  geom_vline(xintercept = best.cutoff.ridge) +
  theme_bw()

```
Predictions for negative sequences


```{r, fig.height = 5, fig.width = 7}

ggplot(dat.neg, aes(x = Ridge_positive_prob, y = Attention_positive_prob)) +
  geom_point(alpha = 0.02) + geom_hline(yintercept = best.cutoff.nn) +
  geom_vline(xintercept = best.cutoff.ridge) +
  theme_bw()

```


# New library

For the new library, we can include 

* 5000 correctly classified positive & negative sequences
* all mis-classified positive sequences (by
either model): `r sum(dat.pos$Ridge_positive_prob < best.cutoff.ridge | dat.pos$Attention_positive_prob < best.cutoff.nn)` sequences
* all negative sequences mis-classified by NN only: `r sum(dat.neg$Attention_positive_prob > best.cutoff.nn & dat.neg$Ridge_positive_prob < best.cutoff.ridge)` sequences
* 5000 sequences mis-classified by ridge regression only
* 5000 sequences mis-classified by both methods


# Motif search

We'll look at motifs discriminating positive/negative sequences that are found
in one method, but not the other.

```{r}

seqs.save <- list(tp.nn = dat.pos$sequence[dat.pos$Attention_positive_prob > best.cutoff.nn &
                                             dat.pos$Ridge_positive_prob < best.cutoff.ridge],
                  tp.ridge = dat.pos$sequence[dat.pos$Attention_positive_prob < best.cutoff.nn &
                                             dat.pos$Ridge_positive_prob > best.cutoff.ridge],
                  tp.both = dat.pos$sequence[dat.pos$Attention_positive_prob > best.cutoff.nn &
                                             dat.pos$Ridge_positive_prob > best.cutoff.ridge],
                  tp.none = dat.pos$sequence[dat.pos$Attention_positive_prob < best.cutoff.nn &
                                             dat.pos$Ridge_positive_prob < best.cutoff.ridge],
                  tn.nn = dat.neg$sequence[dat.neg$Attention_positive_prob < best.cutoff.nn &
                                             dat.neg$Ridge_positive_prob > best.cutoff.ridge],
                  tn.ridge = dat.neg$sequence[dat.neg$Attention_positive_prob > best.cutoff.nn &
                                             dat.neg$Ridge_positive_prob < best.cutoff.ridge],
                  tn.both = dat.neg$sequence[dat.neg$Attention_positive_prob < best.cutoff.nn &
                                             dat.neg$Ridge_positive_prob < best.cutoff.ridge],
                  tn.none = dat.neg$sequence[dat.neg$Attention_positive_prob > best.cutoff.nn &
                                             dat.neg$Ridge_positive_prob > best.cutoff.ridge])

for (i in names(seqs.save)) {
  output.vec <- vector(mode = "character", length = length(seqs.save[[i]]) * 2)
  for (j in 1:length(seqs.save[[i]])) {
    output.vec[j*2 - 1] <- paste0(">seq", j)
    output.vec[j*2] <- seqs.save[[i]][j]
  }
  write.table(output.vec, paste0("../data/xiao_analysis/", i, "-seqs.txt"), 
              sep = "\t", quote = FALSE, row.names = FALSE, col.names = FALSE)
}
```


Generate weblogos (not looking great)


```{r, eval = FALSE}

library(RWebLogo)
options('python_cmd' = "C:/Users/cclass/Anaconda3/pkgs/python-3.7.6-h60c2a47_2")

weblogo(seqs = seqs.save$tp.both)

```

Categorize and draw lineplots

```{r}

library(plyr)
library(tidyr)

prop.mat <- list()
alphabet <- data.frame(AA = factor(c("R", "H", "K", "D", "E", "S", "T", "N", "Q", "C", "G", "P", "A", "V", "I", "L", "M", "F", "Y", "W"),
                                   levels = c("R", "H", "K", "D", "E", "S", "T", "N", "Q", "C", "G", "P", "A", "V", "I", "L", "M", "F", "Y", "W")),
                       type = factor(c(rep("Basic", 3), rep("Acidic", 2), rep("Polar", 4), rep("Special", 3), rep("Aliphatic", 5), rep("Aromatic", 3)),
                                levels = c("Aromatic", "Acidic", "Basic", "Aliphatic", "Polar", "Special")))

for (i in names(seqs.save)) {
  print(i)
  
  seqs.split <- strsplit(seqs.save[[i]], split = "")
  
  prop.tabs <- lapply(1:length(seqs.split[[1]]), 
                      function(j) table(factor(sapply(seqs.split, `[[`, j), levels = levels(alphabet$AA))) / 
                        length(seqs.split))
  
  prop.mat[[i]] <- do.call(cbind, prop.tabs)
  
}

```

Plot by AA

```{r}

prop.tidy <- list()


for (i in names(prop.mat)) {
  prop.df <- cbind(as.data.frame(prop.mat[[i]]), alphabet)
  prop.df$AA <- rownames(prop.df)
  prop.tidy[[i]] <- gather(prop.df, key = "position", value = "frequency", -c("AA", "type"))
  prop.tidy[[i]]$position <- as.numeric(substr(prop.tidy[[i]]$position, 
                                               start = 2, stop = nchar(prop.tidy[[i]]$position)))
}


```

```{r, fig.height = 6, fig.width = 8}

ggplot(prop.tidy$tp.both, aes(x = position, y = frequency, colour = AA)) +
  geom_smooth(se = FALSE) + theme_bw() + facet_wrap(~ type) + 
  geom_hline(yintercept = 1/20, linetype = "dashed") +
  ylab("Frequency of AA")



#ggplot(prop.tidy$tn.nn, aes(x = position, y = frequency, colour = AA)) +
#  geom_line() + theme_bw() + facet_wrap(~ type)

```


```{r, fig.height = 6, fig.width = 8}

ggplot(prop.tidy$tp.nn, aes(x = position, y = frequency, colour = AA)) +
  geom_smooth(se = FALSE) + theme_bw() + facet_wrap(~ type) + 
  geom_hline(yintercept = 1/20, linetype = "dashed") +
  ylab("Frequency of AA")

```


```{r, fig.height = 6, fig.width = 8}

ggplot(prop.tidy$tp.ridge, aes(x = position, y = frequency, colour = AA)) +
  geom_smooth(se = FALSE) + theme_bw() + facet_wrap(~ type) + 
  geom_hline(yintercept = 1/20, linetype = "dashed") +
  ylab("Frequency of AA")

```

```{r, fig.height = 6, fig.width = 8}

ggplot(prop.tidy$tp.none, aes(x = position, y = frequency, colour = AA)) +
  geom_smooth(se = FALSE) + theme_bw() + facet_wrap(~ type) + 
  geom_hline(yintercept = 1/20, linetype = "dashed") +
  ylab("Frequency of AA")

```

```{r, fig.height = 6, fig.width = 8}

ggplot(prop.tidy$tn.both, aes(x = position, y = frequency, colour = AA)) +
  geom_smooth(se = FALSE) + theme_bw() + facet_wrap(~ type) + 
  geom_hline(yintercept = 1/20, linetype = "dashed") +
  ylab("Frequency of AA")


```


```{r, fig.height = 6, fig.width = 8}

ggplot(prop.tidy$tn.nn, aes(x = position, y = frequency, colour = AA)) +
  geom_smooth(se = FALSE) + theme_bw() + facet_wrap(~ type) + 
  geom_hline(yintercept = 1/20, linetype = "dashed") +
  ylab("Frequency of AA")

```


```{r, fig.height = 6, fig.width = 8}

ggplot(prop.tidy$tn.ridge, aes(x = position, y = frequency, colour = AA)) +
  geom_smooth(se = FALSE) + theme_bw() + facet_wrap(~ type) + 
  geom_hline(yintercept = 1/20, linetype = "dashed") +
  ylab("Frequency of AA")

```

```{r, fig.height = 6, fig.width = 8}

ggplot(prop.tidy$tn.none, aes(x = position, y = frequency, colour = AA)) +
  geom_smooth(se = FALSE) + theme_bw() + facet_wrap(~ type) + 
  geom_hline(yintercept = 1/20, linetype = "dashed") +
  ylab("Frequency of AA")

```



## Draft figure for paper

```{r}

for (i in names(prop.tidy)) {
  prop.tidy[[i]]$seqclass <- gsub("\\..*", "", i)
  prop.tidy[[i]]$Correct <- gsub(".*\\.", "", i)
}

# Brad colors? Make distinct

prop.all <- do.call(rbind, prop.tidy)
prop.all$Correct <- factor(prop.all$Correct, levels = c("nn", "both", "ridge", "none"))
levels(prop.all$Correct)[1:2] <- c("LSTM Network", "Both")
prop.all$seqclass <- factor(prop.all$seqclass, levels = c("tp", "tn"))
levels(prop.all$seqclass) <- c("Functional", "Non-Functional")

prop.all$AA <- factor(prop.all$AA, levels = c("W", "F", "Y", "D", "E", "R", "K", "H"))

ggplot(prop.all[prop.all$Correct %in% c("LSTM Network", "Both") &
                  prop.all$type %in% c("Aromatic", "Acidic", "Basic"),], 
       aes(x = position, y = frequency, colour = AA)) +
  stat_smooth(geom = "line", aes(size = Correct, linetype = Correct), se = FALSE) +
  theme_bw() + facet_grid(seqclass ~ type) + 
  ylab("Frequency of AA") +
  #scale_alpha_manual(values = c(1, 0.6)) + 
  scale_size_manual(values = c(1.2, 1)) + 
  scale_linetype_manual(values = c("solid", "dashed")) +
  scale_colour_brewer(palette = "Set1") +
  theme(legend.key.width = unit(1,"cm"))
ggsave("../output/NN-correct-AAfreqs.tiff",
       height = 4, width = 7)

```


```{r}

ggplot(prop.all[prop.all$Correct %in% c("LSTM Network", "Both") &
                  prop.all$type %in% c("Aromatic", "Acidic", "Basic"),], 
       aes(x = position, y = frequency, colour = AA)) +
  stat_smooth(geom = "line", aes(size = Correct, linetype = Correct), se = FALSE) +
  theme_bw() + facet_grid(seqclass ~ type) + 
  ylab("Frequency of AA") +
  #scale_alpha_manual(values = c(1, 0.6)) + 
  scale_size_manual(values = c(1.2, 1)) + 
  scale_linetype_manual(values = c("solid", "dashed")) +
  scale_colour_manual(values = c("#ad7f00", "#ffbb00", "#fcda7c",
                                 "#ba0000", "#ff7070",
                                 "#13008f", "#2200ff", "#9686fc")) +
  theme(legend.key.width = unit(1,"cm"))
ggsave("../output/NN-correct-AAfreqs-cols.tiff",
       height = 4, width = 7)
```
